Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed Evolutionary Computation

### 2015-11-16 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Changed figure and caption
Closes #46
JavaScript considerations

\begin{figure}[!t]
\centering
\includegraphics[width=12cm]{baseline-times.png}
\caption{Comparison of time to solution for the baseline system, using a {\tt
    Node.js} client, for different population sizes. Only
runs in which solution has been found are used for this graph.}
\label{fig:baseline}
\end{figure}
First, we establish the baseline performance by running the
evolutionary algorithm in a desktop client written using NodEO
\cite{nodeo2014}, the basic JavaScript library we have used as base to
build {\sf NodIO}, the framework. This experiment tries to 
find the solution to the 40-trap function with parameters $l=4, a=1,
b=2, z=3$ and a population size equal to 512. The algorithm in a
single island was run until the solution (a string with all
ones) was found or five million evaluations had been performed. It
took around a minute, on average, that is $68.9694$ seconds, to
perform the fifty runs. In this experiment only 33, that is, 66\% were
successful. The experiments were repeated for population $p=1024$,
with success rate upgraded to 100\% and average duration $3.46$
seconds. Results for these two experiments
are shown in Figure \ref{fig:baseline}. The baseline system has these
characteristics: \begin{verbatim}
Linux penny 3.13.0-34-generic #60-Ubuntu SMP Wed Aug 13 15:45:27 UTC
2014 x86_64 x86_64 x86_64 GNU/Linux
\end{verbatim}
with an 8-core {\tt Intel(R) Core(TM) i7-4770 CPU @ 3.40GHz}.

@@

\begin{itemize}
\item {\em Randomize} Both Malab and Java implementations rely on a 
Java Randomizer library as a standardized implementation of the 
randomization of shift vectors and matrices used for the experiments.
The library included the generation of pseudorandom Gaussian of type
double and uniformly distributed integers. Generators used a single
{\tt long} seed. In order to implement the same functionality in
JavaScript, the library random-js was used, as there are inconsistencies 
in the implementation of {\tt Math.random()} between engines in the
number of bits used in the generation and results are non-deterministic. 
The library random-js, implements a Mersenne Twister algorithm producing 
consistent results across all JavaScript implementations assuming the same seed.
\item {\em Arrays} 
\item {\em Timing functions}
\item {\em Tools} 
\end{itemize}

The performance of the time required for 10,000 FEs is presented in Figure \ref{fig:f15_times}.
The test was run on an 3.7 GHz Quad-Core Intel Xeon E5 processor running OS X 10.10.5, 
using Java(TM) SE Runtime Environment (build 1.8.0\_25-b17), 
MatLab version R2015a for Mac OSX, Node FrameWork version 0.12.2, Google Chrome Version 46.0.2490.86 (64-bit) 
and Firefox version 41.0.2.

>>>>

\begin{figure}[!t]
\centering
\includegraphics[width=12cm]{baseline-times.png}
\caption{Comparison of times to solution (labeled as {\bf times}) for the baseline system, using a {\tt
    Node.js} client, for different population sizes. Only
runs in which solution has been found are used for this graph.}
% Plotted with: ~/Code/splash-volunteer/data/plot-baseline.R
\label{fig:baseline}
\end{figure}
First, we establish the baseline performance by running the
evolutionary algorithm in a desktop client written using NodEO
\cite{nodeo2014}, the basic JavaScript library we have used as base to
build {\sf NodIO}, the framework. This experiment tries to 
find the solution to the 40-trap function with parameters $l=4, a=1,
b=2, z=3$ and a population size equal to 512. The algorithm in a
single island was run until the solution (a string with all
ones) was found or five million evaluations had been performed. It
took around a minute, on average, that is $68.9694$ seconds, to
perform the fifty runs. In this experiment only 33, that is, 66\% were
successful. The experiments were repeated for population $p=1024$,
with success rate upgraded to 100\% and average duration $3.46$
seconds. Results for these two experiments
are shown in Figure \ref{fig:baseline}. The baseline system has these
characteristics: \begin{verbatim}
Linux penny 3.13.0-34-generic #60-Ubuntu SMP Wed Aug 13 15:45:27 UTC
2014 x86_64 x86_64 x86_64 GNU/Linux
\end{verbatim}
with an 8-core {\tt Intel(R) Core(TM) i7-4770 CPU @ 3.40GHz}.

@@

\begin{itemize}
\item {\em Randomize} Both Malab and Java implementations rely on a 
Java Randomizer library as a standardized implementation of the 
randomization of shift vectors and matrices used for the experiments.
The library included the generation of pseudorandom Gaussian of type
double and uniformly distributed integers. Generators used a single
{\tt long} seed. In order to implement the same functionality in
JavaScript, the library random-js was used, as there are inconsistencies 
in the implementation of {\tt Math.random()} between engines in the
number of bits used in the generation and results are non-deterministic. 
The library random-js, implements a Mersenne Twister algorithm producing 
consistent results across all JavaScript implementations assuming the same seed.
\item {\em Timing functions} In order to compare the performance of 
each implementation, an accurate and consistent way of measuring 
the runtime is needed. In JavaScript the {\tt Date class} is often used 
to measure execution time, but the maximum resolution is given in miliseconds.
In order to measure the runtime intervals with higher precission two functions
where used: for Node.js we used the native {\tt process.hrtime()} function
which returns a high-resolution measure in a {\tt [seconds, nanoseconds]} array.
For testing in the browser the {\tt Performance.now()} function was used, this
function uses floating-point numbers with up to microsecond precision. 
A drawback found is that the {\tt Performance.now()} function is only implemented
in Firefox and Chrome browsers. Both functions are independent of the system clock. 
\item {\em Tools} An important consideration when chosing certain language to 
to write non-trivial programs is the availability of development tools. When developing
the experiments, we used a comprehensive set of developer tools available in both
browsers and desktop. These tools included package managers, debbugers, logs, 
network monitors, editors and even debugging of the multi-threaded execution of Web Workers.
A drawback found is that developer tools very from browser to browser, 
but they all have the same functionality.
\item {\em Data Types} JavaScript uses floating point numbers with a 
limited precision of 64 bits that partialy implements the functionality of
the {\tt StrictMath} library used by the Java implemantation. If more precision
is needed developers could use {\tt math.js}, an extensive math library with support for matrices, big and complex numbers.

\end{itemize}

We found that the JavaScript ecosystem is mature enough for developing non-trivial 
code, but developers have to consider the differences betwwen VMs implementations
and use certain libraries in order to increase the precision and repeatability of
algorithms. To test the performance of JavaScript against Java and Matlab

The test was run on an 3.7 GHz Quad-Core Intel Xeon E5 processor running OS X 10.10.5, 
using Java(TM) SE Runtime Environment (build 1.8.0\_25-b17), 
MatLab version R2015a for Mac OSX, Node FrameWork version 0.12.2, Google Chrome Version 46.0.2490.86 (64-bit) 
and Firefox version 41.0.2.  The performance of the time required for 10,000 FEs is presented in Figure \ref{fig:f15_times}.
